---
title: "Fee_data"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Adrian D. Garcia"
date: "June 2020"
output: html_document
---

```{r setup, message = FALSE}
library(tidyverse) #For Data Analysis
library(lubridate) #For working with dates
library(DT) #For visualizing tables
```

First, I pull in the data from Morningstar Direct (combined using “Combine_Files.r”). There are 25,616 rows and 88 columns. Data was pulled from Morningstar Direct on June 25, 2020.


```{r load, message = FALSE}

Full <- read_csv("MS_Fee_Report_062520_Full.csv",
                 guess_max = 25000)

```

I’m making the following adjustment to the data:

-Replaces spaces in the column names with periods.
-Calculate the average assets for 2017, 2018 and 2019.
-Multipy the annual average asset figures by the expense ratio (to use later for asset-weighted calculations)
-Calculate the percent change in net expense ratios from 2017 to 2018 and 2018 to 2019. 

```{r adjustment, Message=FALSE}

Adjusted_Data <- Full %>% 
  rename_all(make.names) %>% 
  mutate(Net_Asst_2017_AVG = rowMeans(.[,53:64], na.rm = TRUE),
         Net_Asst_2018_AVG = rowMeans(.[,65:76], na.rm = TRUE),
         Net_Asst_2019_AVG = rowMeans(.[,65:76], na.rm = TRUE),
         Asst_By_Fee_2017 = Net_Asst_2017_AVG * Annual.Report..Net.Expense..Ratio..Year2017,
         Asst_By_Fee_2018 = Net_Asst_2017_AVG * Annual.Report..Net.Expense..Ratio..Year2018,
         Asst_By_Fee_2019 = Net_Asst_2017_AVG * Annual.Report..Net.Expense..Ratio..Year2019,
         Pct_Change_2017_2018 = (Annual.Report..Net.Expense..Ratio..Year2018 - Annual.Report..Net.Expense..Ratio..Year2017)/Annual.Report..Net.Expense..Ratio..Year2017 *100,
         Pct_Change_2018_2019 = (Annual.Report..Net.Expense..Ratio..Year2019 - Annual.Report..Net.Expense..Ratio..Year2018)/Annual.Report..Net.Expense..Ratio..Year2018 *100)
  
write_csv(Adjusted_Data, "MS_Fee_Report_062520_ag.csv")

```

Here I want to look how deep firms slashed pay ratios from 2018 to 2019. First, remove rows with an expense ratio of zero in 2018 because I can’t calculate changes with that base. I also filter out expense ratios listed as NA (blank) for 2018 and 2019.

Now I need to put the percent change in fees from 2018 to 2019 into buckets. I did that with the cut() function. Here’s a for personal reference<https://rpubs.com/pierrelafortune/cutdocumentation>.

```{r category, Message=FALSE}
Category <- Adjusted_Data %>% 
  select(
    1:20,
    86:96
  ) %>%
  filter(
    Annual.Report..Net.Expense..Ratio..Year2018 != 0 &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2018) &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2019)
  ) %>% 
  mutate(
    FeeChange = 
      cut(Pct_Change_2018_2019,
          breaks = c(-Inf,-50,-25,-10,-1,-0.01,0,Inf),
          labels = c("< -50", "-25 - -50", "-10 - -25", "-1 - 10","0 - -1", "0", "> 0"),
          include.lowest = TRUE,
          right = TRUE
      )
  )

write_csv(Category,"shareclass_cuts_total.csv")

Summary_Category <- Category %>% 
  group_by(FeeChange) %>% 
  summarise(count = n())

ggplot(data = Summary_Category, mapping = aes(x = FeeChange, y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = 0)
```

Here I am looking at the fund families that with the lowest and highest overall fees across active products. I start by filtering out index funds and removing fees that are NA (blank) for 2019. Next I group the data by firm, sum their adjusted assets, sum their net assets and divide the total adjusted data by the total net asssets to calculate asset-weighted fees for each fund shop. I select the top and bottom 15 fund families based on the results.

```{r active_funds, Message=FALSE}
ActiveFunds <- Adjusted_Data %>% 
  filter(
    Net_Asst_2019_AVG != 0 &
   Index..Fund != "Yes" &
     !is.na(Annual.Report..Net.Expense..Ratio..Year2019)
  )

Active_Fees <- ActiveFunds %>% 
  group_by(Firm.Name) %>% 
  summarise(
    Sum_Assts_Wted = sum(Asst_By_Fee_2019, na.rm = TRUE),
    Sum_Assets = sum(Net_Asst_2019_AVG, na.rm = TRUE),
    Fee_Wted = Sum_Assts_Wted/Sum_Assets
  )

write_csv(Active_Fees, "fees_activefunds_firm.csv")

Active_HighFees <- Active_Fees %>% 
  arrange(desc(Fee_Wted)) %>% 
  top_n(15, Fee_Wted)

Active_LowFees <- Active_Fees %>% 
  arrange(Fee_Wted) %>% 
  top_n(-15, Fee_Wted)

ggplot(data = Active_HighFees, mapping = aes(x = reorder(Firm.Name, Fee_Wted), y = Fee_Wted)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  geom_text(aes(label = round(Fee_Wted,2)), hjust = 0)

ggplot(data = Active_LowFees, mapping = aes(x = reorder(Firm.Name, -Fee_Wted), y = Fee_Wted)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  geom_text(aes(label = round(Fee_Wted,2)), hjust = 0)
```

Here I am looking at the fund families that with the lowest and highest overall fees across passive products. I start by filtering out non-index funds and removing fees that are NA (blank) for 2019. Next I group the data by firm, sum their adjusted assets, sum their net assets and divide the total adjusted data by the total net asssets to calculate asset-weighted fees for each fund shop. I select the top and bottom 15 fund families based on the results.

```{r passive_funds, Message=FALSE}
PassiveFunds <- Adjusted_Data %>% 
  filter(
    Net_Asst_2019_AVG != 0 &
   Index..Fund != "No" &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2019)
  )

Passive_Fees <- PassiveFunds %>% 
  group_by(Firm.Name) %>% 
  summarise(
    Sum_Assts_Wted = sum(Asst_By_Fee_2019, na.rm = TRUE),
    Sum_Assets = sum(Net_Asst_2019_AVG, na.rm = TRUE),
    Fee_Wted = Sum_Assts_Wted/Sum_Assets
  )

write_csv(Passive_Fees, "fees_passivefunds_firm.csv")


Passive_HighFees <- Passive_Fees %>% 
  arrange(desc(Fee_Wted)) %>% 
  top_n(15)

Passive_LowFees <- Passive_Fees %>% 
  arrange(Fee_Wted) %>% 
  top_n(-15)

ggplot(data = Passive_HighFees, mapping = aes(x = reorder(Firm.Name, -Fee_Wted), y = Fee_Wted)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  geom_text(aes(label = round(Fee_Wted,2)), hjust = 0)

ggplot(data = Passive_LowFees, mapping = aes(x = reorder(Firm.Name, -Fee_Wted), y = Fee_Wted)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  geom_text(aes(label = round(Fee_Wted,2)), hjust = 0)
```

Here I am looking at how many share classes had fees that were cut and hiked as well as saw no change. I start by filtering out fees that are NA (blank) for 2019 and 2018. I also remove fees with a zero expense ratio for 2018 because we can’t calculate change with that base. Next I create a category (FeeChange) that puts the percent changes (continous) in buckets (categorical). After that I just count the buckets.

```{r shareclass, Message=FALSE}
Shareclass <- Adjusted_Data %>% 
  filter(
    Annual.Report..Net.Expense..Ratio..Year2018 != 0 &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2018) &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2019)
  ) %>% 
  mutate(
    FeeChange = 
      cut(Pct_Change_2018_2019,
          breaks = c(-Inf,-0.001,0,Inf),
          labels = c("Cut", "No Change", "Hike"),
          include.lowest = TRUE,
          right = TRUE
      )
  )

write_csv(Shareclass, "shareclass_fee_changes.csv")

Summary_Shareclass <- Shareclass %>% 
  group_by(Firm.Name,
           FeeChange) %>% 
  summarise(count = n())

ggplot(data = Summary_Shareclass, mapping = aes(x = FeeChange, y = count)) +
  geom_bar(stat = "identity")
```

Here I am looking at how many Passive share classes had fees that were cut and hiked as well as saw no change. I start by filtering out non-index funds and fees that are NA (blank) for 2019 and 2018. I also remove fees with a zero expense ratio for 2018 because we can’t calculate change with that base. Next I create a category (FeeChange) that puts the percent changes (continous) in buckets (categorical). After that I just count the buckets.

```{r passive_shareclass, Message=FALSE}
Passive_Shareclass <- Adjusted_Data %>% 
  filter(
    Annual.Report..Net.Expense..Ratio..Year2018 != 0 &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2018) &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2019) &
     Index..Fund == "Yes"
  ) %>% 
  mutate(
    FeeChange = 
      cut(Pct_Change_2018_2019,
          breaks = c(-Inf,-0.001,0,Inf),
          labels = c("Cut", "No Change", "Hike"),
          include.lowest = TRUE,
          right = TRUE
      )
  )

Passive_Summary_Shareclass <- Passive_Shareclass %>% 
  group_by(Firm.Name,
           FeeChange) %>% 
  summarise(count = n()) %>% 
  group_by(Firm.Name) %>% 
  mutate(
    pct = count/sum(count)*100)

write_csv(Passive_Summary_Shareclass, "shareclass_changes_passivefunds.csv")

Passive_Summary_Shareclass %>% 
  filter(FeeChange == "Cut") %>% 
  arrange(desc(count)) %>% 
  ungroup() %>% 
  top_n(15, count) %>% 
ggplot(mapping = aes(x = reorder(Firm.Name, count), y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = count), vjust = 0)
```

tk

```{r active_shareclass, Message = FALSE}

Active_Shareclass <- Adjusted_Data %>% 
  filter(
    Annual.Report..Net.Expense..Ratio..Year2018 != 0 &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2018) &
      !is.na(Annual.Report..Net.Expense..Ratio..Year2019) &
     Index..Fund == "Yes"
  ) %>% 
  mutate(
    FeeChange = 
      cut(Pct_Change_2018_2019,
          breaks = c(-Inf,-0.001,0,Inf),
          labels = c("Cut", "No Change", "Hike"),
          include.lowest = TRUE,
          right = TRUE
      )
  )

Active_Summary_Shareclass <- Active_Shareclass %>% 
  group_by(Firm.Name,
           FeeChange) %>% 
  summarise(count = n()) %>% 
  group_by(Firm.Name) %>% 
  mutate(
    pct = count/sum(count)*100)

write_csv(Active_Summary_Shareclass, "shareclass_changes_activefunds.csv")


Active_Summary_Shareclass %>% 
  filter(FeeChange == "Cut") %>% 
  arrange(desc(count)) %>% 
  ungroup() %>% 
  top_n(15, count) %>% 
ggplot(mapping = aes(x = reorder(Firm.Name, count), y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = count), vjust = 0)

```